---
title: "Reproducible Reporting in R using Quarto"
format: html
editor: visual
---

# Introduction

This is a quarto practical using the **Mortality** and [**Gapminder**](https://www.gapminder.org/) dataset.

## Installing and loading packages

### Installing packages

For packages, it is important to set the CRAN mirror to avoid issues during installation, and it's a good practice to install packages before rendering.

```{}
```

```{r}
#| label: install-packages
#| include: false  ## to avoid the code chunk in the output
options(repos = c(CRAN = "https://cloud.r-project.org"))
install.packages(c("skimr", "gapminder", "plotly", "gt"))
```

### Loading packages

```{r}
#| label: load-packages
#| message: false
#| warning: false
library(tidyverse)
library(haven)
library(skimr)
library(gapminder)
library(gt)
library(plotly)
```

## Import dataset

```{r}
mortality <- read_dta("Data/mortality17.dta")
mortality <- as_factor(mortality) ## To see it as categorical variables
glimpse(mortality) 
```

### Skim the dataset using the "skimr" package

```{r}
skim(mortality)
```

# Data visualization

## Table

The table below shows deaths at a younger age with higher diastolic blood pressure from the mortality dataset.

```{r}
#| label: death-at-younger-dbp
#| tbl-cap: Deaths at younger age


```

## Figure

The figure below shows a scatter plot of systolic blood pressure by age.

```{r}
#| label: SBP-by-age
#| fig-cap: A scatter plot of systolic blood pressure by age
ggplot(mortality, aes(x = age, y = systolic)) +
  geom_point(aes(color = sex), alpha=0.7) +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(
    title = "Blood Pressure by Sex (Color) and Area (Shape)",
    x = "Age in years", y = "Systolic (mmHg)")
```

## Using the Gapminder data

Gapminder combines data from multiple sources into unique, coherent time-series that can’t be found elsewhere.[@bryan2015]

The table below displays the Gapminder data of Ethiopia.

```{r}
#| label: Gapminder-Ethiopia
#| tbl-cap: Gapminder data of Ethiopia (1952 - 2007)
gapminder <- gapminder
glimpse(gapminder)
gapminder |> 
  filter(country == "Ethiopia") |> 
  gt()
```

### Animated visualisation

```{r}
#| label: gapminder-visualisation
#| fig-cap: Life expectancy by GDP per Capita from 1952 - 2007 

A <- ggplot(gapminder, aes(gdpPercap, lifeExp, color = continent)) +
  geom_point(aes(size = pop, frame = year, ids = country)) + 
  scale_x_log10()
#A
An_fig <- ggplotly(A)

An_fig <- An_fig |>  
  animation_opts(
    1000, easing = "linear", redraw = FALSE)

An_fig
```
